Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-06-14 manuscript.Rmd
typos

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different measurements. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020]. For other metrics, such as the number of "conversational turns" between the child and communicative partners and the identification of male speakers, agreement was poor. Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` sessions from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` sessions from older infants. Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`).

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019]. Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest, to determine whether data could be simultaneously recorded from the LENA&reg; and IMU sensors. However, LENA&reg; data were not analyzed as part of the current study.

>>>>

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA&reg; algorithms found mixed results about the accuracy of different measurements. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA&reg;'s automatic counts were strong, *r* = .698 and *r* = .649, respectively [@CristiaLavechin2020]. For other metrics, such as the number of "conversational turns" between the child and communicative partners, agreement was poor (*r* = .364). Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Most participants were tested in a single session (*n* = `r nrow(num_sessions %>% filter(n == 1))`), but `r nrow(num_sessions %>% filter(n > 1))` participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of `r length(unique(ppts$id_uni))` sessions, with `r num_sessions %>% filter(group == "Younger") %>% pull(n) %>% sum` sessions from younger infants and `r num_sessions %>% filter(group == "Older") %>% pull(n) %>% sum` sessions from older infants. Across sessions, younger infants' age ranged from `r apa_num(age["Younger","min"], digits = 1)` to `r apa_num(age["Younger","max"], digits = 1)` months (*M* = `r apa_num(age["Younger","mean"], digits = 1)`) and older infants' age ranged from `r apa_num(age["Older","min"], digits = 1)` to `r apa_num(age["Older","max"], digits = 1)` months (*M* = `r apa_num(age["Older","mean"], digits = 1)`). The total number of recording sessions (`r length(unique(ppts$id_uni))`) exceeded the number of sessions employed in comparable past work: 10 in @NamPark2013, 15 in @FranchakScott2021, 22 in @AiraksinenRasanen2020, and 33 in @YaoPlotz2019.

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. We chose a sampling rate of 62.5 Hz (one of the available preset rates) based on prior work that used rates of 50-64 Hz [@FranchakScott2021; @AiraksinenRasanen2020; @YaoPlotz2019]. Infants also wore a LENA&reg; recorder throughout the day in the front pocket of a LENA&reg; shirt, located near the infant’s chest, to determine whether data could be simultaneously recorded from the LENA&reg; and IMU sensors. However, LENA&reg; data were not analyzed as part of the current study.

### 2023-06-14 manuscript.tex
typos

From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings. Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction. When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position (Franchak et al., 2018; Kretch et al., 2014; Luo \& Franchak, 2020). Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration (Soska \& Adolph, 2014). Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position (Chen et al., 2022; Karasik et al., 2011). As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017). Caregivers provide different learning input to infants depending on infants' body position (Karasik et al., 2014; Kretch et al., 2022; West \& Iverson, 2021). For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects. Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning (Franchak, 2020).

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion (Airaksinen et al., 2020) and infant-caregiver contact (Yao et al., 2019). We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).

@@

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are \emph{mobile}. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is \emph{unobtrusive}. Participants' reactivity to observation, such as from a video camera, may influence behavior. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019). Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (Malachowski et al., 2023; e.g., Wass et al., 2022). Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018). ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021). Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA® algorithms found mixed results about the accuracy of different measurements. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020). For other metrics, such as the number of ``conversational turns'' between the child and communicative partners and the identification of male speakers, agreement was poor. Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate. First, the magnitude of movement can vary greatly \emph{within} a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways (Adolph et al., 1998). Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, or in a ``W'' position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso. Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity (Kwon et al., 2019; Patel et al., 2019). Although Greenspan et al. (2021) were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground. Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults (Arif \& Kattan, 2015; Preece et al., 2009), children (Nam \& Park, 2013; Ren et al., 2016; Stewart et al., 2018), and infants (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019). Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated. Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotation.

@@

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7).

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data at 62.5 Hz throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest.

Video recordings were captured using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording. As described below, the video recording was not continuous. The camera created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).

@@

Because the camera only had the battery life to record for \textasciitilde3 hours (split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was automatically recorded, termed the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because there was a variable gap of \textasciitilde40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

We report three sets of results based on 34 full-day testing sessions resulting in a total of 302 hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}). The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.04 hours. Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.25 hours. Body position data were available to describe 37.3-100.0\% of the awake portion of the recording period (median = 100.0\%). Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of 6.14 hours of motion data, with the entire dataset totaling 206.66 hours. Even the shortest full recording (3.06 hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

Second, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data. Although most caregivers were diligent about completing logs, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4). In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body. Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on hips/ankles. Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.

Third, although the goal was to collect fully naturalistic behavior, we opted to restrict recording to times when participants were in the home. Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands. However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions. Future work should aim to collect data in these scenarios so that body position can be captured in or out of the home.

>>>>

From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings. Changes in \emph{body position}---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction. When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position (Franchak et al., 2018; Kretch et al., 2014; Luo \& Franchak, 2020). Learning to sit facilitates object exploration (Soska \& Adolph, 2014). Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position (Chen et al., 2022; Karasik et al., 2011). As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone (Adolph \& Tamis-LeMonda, 2014; Franchak et al., 2018; Franchak, 2019; Thurman \& Corbetta, 2017). Caregivers provide different learning input to infants depending on infants' body position (Karasik et al., 2014; Kretch et al., 2022; West \& Iverson, 2021). For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects. Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning (Franchak, 2020).

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions (Airaksinen et al., 2022, 2020; Franchak et al., 2021), the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a \emph{distal comparison}.

@@

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are \emph{mobile}. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is more \emph{unobtrusive}. Participants' reactivity to observation, such as from a video camera, may influence behavior more compared with a sensor that records only motion or audio data. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA® device (Bergelson, Amatuni, et al., 2019). Third and fourth, recordings capture \emph{real-time data} over a \emph{full day}. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual (Malachowski et al., 2023; e.g., Wass et al., 2022). Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) (Kadooka et al., 2021, April; Tamis-LeMonda et al., 2018). ``Burstiness'' of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent (Barbaro \& Fausey, 2022; Warlaumont et al., 2021). Fifth, \emph{automatic classification} means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

But automatic classification can only replace human annotation if it is sufficiently accurate and unbiased. An independent assessment of the LENA® algorithms found mixed results about the accuracy of different measurements. For example, correlations between human transcribed counts of adult words and child vocalizations against LENA®'s automatic counts were strong, \emph{r} = .698 and \emph{r} = .649, respectively (Cristia et al., 2020). For other metrics, such as the number of ``conversational turns'' between the child and communicative partners, agreement was poor (\emph{r} = .364). Thus, for some use cases (and for some metrics), long form audio recordings provide a mobile, unobtrusive way to automatically score real-time data over a full day. In the remainder of the paper, we turn to the question of how to replicate those key features in long form recordings of motor behavior.

@@

Categorizing body position---supine on the back, prone on the belly, sitting, upright, or held off the ground by a caregiver---is too complex for cut-point definitions to be accurate. First, the magnitude of movement can vary greatly \emph{within} a body position. An upright infant can be standing still or can be walking briskly across the room. A prone infant can be stationary in ``tummy time'', or they can crawl in a myriad of ways (Adolph et al., 1998). Moreover, the configuration of the arms, legs, and torso within a body position can vary greatly in everyday contexts. Infants can sit on the floor in a tripod position with support from an arm, in a ``V'' position with legs fully extended, or in a ``W'' position with knees bent. Sitting on a caregiver's lap without the need to maintain balance means that the legs can dangle and the torso can lean in different directions. Sitting in a high chair or car seat reduces the magnitude of postural sway within sitting, and creates even more possibilities for how the arms and legs may move relative to the torso. Indeed, creating an all-encompassing set of rules for how to annotate sitting from video is no trivial task because of the various ways that sitting can occur in daily life. Finally, caregivers frequently pick up and transport infants, creating motion signals that need to be differentiated from independent activity (Kwon et al., 2019; Patel et al., 2019). Although Greenspan et al. (2021) were able to use cut-point definitions about infant trunk angle to categorize body positions, this method was not able to differentiate times that infants were held by caregivers off the ground. Thus, modern approaches to human activity recognition have used machine learning to classify activity categories based on features derived from IMU data in adults (Arif \& Kattan, 2015; Preece et al., 2009), children (Nam \& Park, 2013; Ren et al., 2016; Stewart et al., 2018), and infants (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019). Synchronized video with ground-truth human annotations creates training data for a machine learning algorithm, such as a random forest model, that can be later used to predict categories from IMU data that was not annotated. Crucially, this allows automatic classification to scale to full day recording by relying on a relatively smaller set of video annotations.

@@

Most participants were tested in a single session (\emph{n} = 15), but 7 participants contributed between 2-4 sessions as part of an ongoing longitudinal study. Although including more data from some participants could over-represent their characteristics in the model, we reasoned that this drawback was outweighed by having more available data to use for training. Only 1 session was excluded due to a technical error---one of the four IMU sensors failed to record, resulting in an unusable set of data for classification. Across the two age groups, we report data on a total of 34 sessions, with 14 sessions from younger infants and 20 sessions from older infants. Across sessions, younger infants' age ranged from 3.8 to 7.2 months (\emph{M} = 5.2) and older infants' age ranged from 10.7 to 14.2 months (\emph{M} = 11.7). The total number of recording sessions (34) exceeded the number of sessions employed in comparable past work: 10 in @Nam and Park (2013), 15 in Franchak et al. (2021), 22 in Airaksinen et al. (2020), and 33 in Yao et al. (2019).

@@

Four inertial movement units (IMUs) were used to record infant movement across the day (MC10 Biostamp). A custom garment was designed to hold the IMUs: Internal pockets were sewn into a snug-fitting pair of infant leggings so that IMUs would stay close to the body (reducing vibration), which also prevented infants from pulling out the sensors. A pocket over the hip and a pocket just above the ankle were sewn on the lateral surface of the right and left legs of the garment. Each sensor recorded accelerometer and gyroscope data throughout the day, with sufficient battery and onboard storage to record for approximately 12 hours. We chose a sampling rate of 62.5 Hz (on of several device presets) based on prior work that used rates of 50-64 Hz (Airaksinen et al., 2020; Franchak et al., 2021; Yao et al., 2019). Infants also wore a LENA® recorder throughout the day in the front pocket of a LENA® shirt, located near the infant's chest, to determine whether data could be simultaneously recorded from the LENA® and IMU sensors. However, LENA® data were not analyzed as part of the current study.

Video recordings were captured at 30 Hz using an action camera on a miniature tripod (Insta360 ONE R) that caregivers placed in the room that the infant was in. The proprietary ``Boosted Battery Base'' attached to the action camera to allow for a total of 3 hours of recording. As described below, the video recording was not continuous. The camera created two video files temporally separated by a gap of approximately 40-45 s. Caregivers also received a log sheet to record times that infants napped as well as times that the sensor garment was removed from the infant (e.g., baths, diaper changes, errands).

@@

Because the camera only had the battery life to record for \textasciitilde3 hours (automatically split into two 90-minute video files), this divided the day into different periods for analysis. As seen in the bottom of Figure \ref{fig:exemplar-timeline}), the \emph{video period} comprised the first three hours of recording starting from the researcher's arrival when they turned on the camera. The first 90-minute video file, termed the \emph{proximal comparison}, contained the activities during the guided call followed by a period during which infants and caregivers resumed their normal activities. Because this video contained the synchronization point, the data in this period had high temporal synchrony between IMU and video data. Synchronization errors were estimated to be less than 30-60 ms (1-2 video frames). When the first video file ended, a second video file was automatically recorded, termed the \emph{distal comparison}. This video recorded the next 90 minutes of natural activity. However, because of a limitation in the camera, there was a variable gap of \textasciitilde40 s between the two videos, synchronization in the distal comparison video was coarser, with estimated temporal offsets of \textasciitilde5 s in either direction.

@@

We report three sets of results based on 34 full-day testing sessions resulting in a total of 302 hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides, to our knowledge, the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

Among participants who wore the sensors from morning to bedtime, all infants wore the sensor garment during 100\% of the intended recording period, excluding the times caregivers were asked to remove the garment (naps, diaper changes, trips out of the house), based on caregiver logs of wear time. No caregiver mentioned needing to remove the garment for any other reason (such as infant discomfort), meaning that the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \ref{fig:timelines}). The total length of the recording period ranged from 6.92-11.75 hours with a median of 9.04 hours. Nap times reported during the recording period ranged from 0.00-4.99 hours with a median of 2.25 hours. Body position data were available to describe 37.3-100.0\% of the awake portion of the recording period (median = 100.0\%). Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of 6.14 hours of motion data, with the entire dataset totaling 206.66 hours. Even the shortest full recording (3.06 hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

Second, we relied on caregiver logs of infants' naps and times when sensors were removed to separate usable data (times when infants wore the sensors while awake) from unusable data. Future work should aim to further assess caregivers' perceptions about the usability and comfort of the garment. Although infants wore the sensor garment throughout all of the desired times, we did not collect independent data (aside from caregivers' logs) to verify wear time. Most caregivers were diligent about completing logs, however, there were a few cases that caregivers may have failed to report naps (younger infants \#1 and \#4). In the future, algorithms can be developed to automatically identify periods of sleep and times when sensors are off the body. Such algorithms exist for adults using wrist-worn sensors, however, we did not use them because they have not been validated for infant participants wearing sensors on hips/ankles. Differences in infants' positioning while asleep (held in caregivers arms, laying in cribs, reclined in strollers and carseats) might make classifying sleep more difficult compared with adults based on movement.

Third, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors. We opted to restrict recording to times when participants were in the home. Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands. However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions. It is also unknown whether participants might have behaved differently had they not worn the sensors. The added bulk in the garment from the sensors (including the LENA® recorder) might have made some positions more uncomfortable, such as lying prone with the LENA® recorder worn on the chest, and/or influenced how caregivers chose to position infants throughout the day.
